---
title: "Session 2"
format: html
---

IDEA: Bayesian power spectrum estimation

# Introduction

```{r}
library("locits")
library("zoo")
library("forecast")
library("tibble")
library("lpacf")
library("forecastLSW")
library("BootWPTOS")
```

```{r}
print(load(url("https://www.ma.imperial.ac.uk/~gnason/Session2.RData")))
```

## Simulated Stationary Time Series

```{r}
set.seed(423)

sim1 <- arima.sim(n=512, model=list(ma=c(0.9, -0.7)))
plot.ts(sim1)
```

```{r}
acf(sim1)
```

## Boris Johnson approval ratings

```{r}
plot(boris.well.zoo, xlab="Date", ylab="Boris Johnson Approval Rating %")
```

```{r}
par(mfrow=c(1,2), pty="s")
acf(coredata(boris.well.zoo), main = "ACF: Boris Well %")
pacf(coredata(boris.well.zoo), main = "PACF: Boris Well %")
```

## Oxford Rain Time Series

```{r}
plot.ts(oxford.rain)
```

```{r}
par(mfrow=c(1,2), pty="s")
acf(oxford.rain, main = "ACF: Oxford Rain")
pacf(oxford.rain, main = "PACF: Oxford Rain")
```

```{r}
stats::spectrum(oxford.rain, spans=c(17,15,13,11))
```

The units are in "oscillations per year", i.e., one up-and-down cycle, so the peak at 1 is a period of 1 year or 12 months.

```{r}
stats::spectrum(oxford.rain, spans=c(30))
```

## BabyECG Series

```{r}
plot.ts(BabyECG)
```

```{r}
par(mfrow=c(1,2), pty="s")
acf(BabyECG[1:256], main = "a.")
acf(BabyECG[1001:1256], main = "b.")
```

## Haar concat simulation

```{r}
set.seed(345)
x <- HaarConcat()
plot.ts(x)
```

```{r}
par(mfrow=c(1,2), pty="s")
acf(x[1:128])
acf(x[358:512])
```

```{r}
hwtos2(x)
```

## Exercise

```{r}
hwtos2(x, filter.number=4)
```

```{r}
plot(hwtos2(x))
```


## BabyECG test of stationarity

```{r}
plot(hwtos2(BabyECG))
```


```{r}
print(hwtos2(BabyECG))
```


## Oxford Rain

```{r}
oxford.rain.extend <- c(rep(mean(oxford.rain), 20), oxford.rain)
hwtos2(oxford.rain.extend)
```


# Wavelet Spectrum Estimation

## Concatenated Haar Process

```{r}
set.seed(345)

y <- HaarConcat()
HC_bw <- AutoBestBW(y)
cat("AutoBestBW is ", HC_bw, "\n")

HC_S <- ewspec3(y, filter.number=1, family="DaubExPhase", binwidth=HC_bw)$S

plot(HC_S)

hc1 <- c(rep(0.5, 128), rep(0, 512-128))
hc2 <- c(rep(0, 128), rep(0.5, 128), rep(0, 256))
hc3 <- rev(hc2)
hc4 <- rev(hc1)

lines(1:512, 1+hc1, lwd=2, col=2)
lines(1:512, 2+hc2, lwd=2, col=2)
lines(1:512, 3+hc3, lwd=2, col=2)
lines(1:512, 4+hc4, lwd=2, col=2)

```


```{r}
HC_S <- ewspec3(y, filter.number=1, family="DaubExPhase", binwidth=100)$S

plot(HC_S)

hc1 <- c(rep(0.5, 128), rep(0, 512-128))
hc2 <- c(rep(0, 128), rep(0.5, 128), rep(0, 256))
hc3 <- rev(hc2)
hc4 <- rev(hc1)

lines(1:512, 1+hc1, lwd=2, col=2)
lines(1:512, 2+hc2, lwd=2, col=2)
lines(1:512, 3+hc3, lwd=2, col=2)
lines(1:512, 4+hc4, lwd=2, col=2)

```


## Exercise

```{r}
#dat <- readRDS("34_ra208.154_dec-63.266_MAXIJ1348TraPDB_andersson.RDS")
dat <- read.csv("Omi_Ceti.csv")

ts <- dat$magnitude
plot.ts(ts)
```

```{r}
length(ts)
```


```{r}
ts_trunc <- ts[1:2^12]

hwtos2(ts_trunc)
```

```{r}
plot(hwtos2(ts_trunc))
```


```{r}
ts_bw <- AutoBestBW(ts_trunc)
cat("AutoBestBW is ", ts_bw, "\n")

ts_spectrum <- ewspec3(ts_trunc, filter.number=1, family="DaubExPhase", binwidth=ts_bw)$S

plot(ts_spectrum)
```

